### 2015-12-01 paper.tex
U_infty is defined in Re def
Fix broken fig reference and trim whitespace

As a cross-flow turbine, the RM2 blades experience large variations in angles of
attack as they rotate about their axis (``cross-flow turbine'' means the axis of
rotation is perpendicular to flow direction---either vertically or
horizontally). These variations become larger as the tip speed ratio
\begin{equation}
\lambda=\frac{\omega R}{U_\infty},
\end{equation}\textbf{\textit{[define Uinf. it looks like a list of notation at the end of the paper could help a alot]}}
where $\omega$ and $R$ are the rotor angular velocity and radius, decreases
\cite{Para2002}. Typically, the maximum angles of attack are sufficiently large
that the blades operate under dynamic stall, which is a complex unsteady process
and deviates significantly from static foil behavior, during part of the
rotation. Furthermore, the higher the solidity of a cross-flow turbine, the
lower the optimal tip-speed ratio at which it operates. Marine Hydrokinetic
(MHK) cross-flow turbines typically have higher solidity than cross-flow wind
turbines (VAWT), and hence operate at lower tip-speed ratios. Since MHK turbines
operate in a higher density fluid, unsteady dynamic effects related to the
blades' pitching motion and flow curvature also become more important when
compared to wind turbines.

@@

Experiments were performed in the UNH tow/wave tank, a 36 m long facility with a
3.66 m wide by 2.44 m deep cross-section, capable of tow speeds up to 3 m/s,
pictured in Figure~\ref{fig:tow-tank}\textbf{\textit{[missing fig?]}}. The turbine was mounted in a frame built
from NACA 0020 struts, attached to the tow carriage by four linear bearings,
which transfer all streamwise force to a pair of S-beam load cells \textbf{\textit{[need reference for each sensor - load cells, ADV, etc.]}}. The turbine
shaft RPM was controlled by a servo motor system, which allowed prescription of
the turbine tip speed ratio. The load torque was measured by an inline rotary
torque transducer. A load cell mounted, mounted at a fixed distance from the servo
motor, provides a redundant measurement. Turbine shaft angle was measured using
the servo drive's emulated encoder output, set to $10^5$ counts per turbine
shaft revolution. Carriage speed, and therefore inflow velocity was measured
using a linear encoder with 10 $\mu$m resolution. All of these
performance-related quantities were sampled at 2 kHz, while the tow tank's
motion controller provided redundant measurements of the carriage speed \textbf{\textit{[need to say how well the agreement between these two measurements are and which one is used throughout the paper; else, remove one of these]}} turbine angular velocity sampled at 1 kHz. Turbine wake measurements at 1 meter
downstream were measured with a Nortek Vectrino+ acoustic Doppler velocimeter,
sampling at 200 Hz. A photo and drawing of the experimental setup are shown in
Fig.~\ref{fig:exp-setup-photo} and Fig.~\ref{fig:exp-setup}, respectively.

>>>>

As a cross-flow turbine, the RM2 blades experience large variations in angles of
attack as they rotate about their axis (``cross-flow turbine'' means the axis of
rotation is perpendicular to flow direction---either vertically or
horizontally). These variations become larger as the tip speed ratio
\begin{equation}
\lambda=\frac{\omega R}{U_\infty},
\end{equation}
where $\omega$ and $R$ are the rotor angular velocity and radius, decreases
\cite{Para2002}. Typically, the maximum angles of attack are sufficiently large
that the blades operate under dynamic stall, which is a complex unsteady process
and deviates significantly from static foil behavior, during part of the
rotation. Furthermore, the higher the solidity of a cross-flow turbine, the
lower the optimal tip-speed ratio at which it operates. Marine Hydrokinetic
(MHK) cross-flow turbines typically have higher solidity than cross-flow wind
turbines (VAWT), and hence operate at lower tip-speed ratios. Since MHK turbines
operate in a higher density fluid, unsteady dynamic effects related to the
blades' pitching motion and flow curvature also become more important when
compared to wind turbines.

@@

Experiments were performed in the UNH tow/wave tank, a 36 m long facility with a
3.66 m wide by 2.44 m deep cross-section, capable of tow speeds up to 3 m/s,
pictured in Fig.~\ref{fig:exp-setup}. The turbine was mounted in a frame built
from NACA 0020 struts, attached to the tow carriage by four linear bearings,
which transfer all streamwise force to a pair of S-beam load cells \textbf{\textit{[need reference for each sensor - load cells, ADV, etc.]}}. The turbine
shaft RPM was controlled by a servo motor system, which allowed prescription of
the turbine tip speed ratio. The load torque was measured by an inline rotary
torque transducer. A load cell mounted, mounted at a fixed distance from the servo
motor, provides a redundant measurement. Turbine shaft angle was measured using
the servo drive's emulated encoder output, set to $10^5$ counts per turbine
shaft revolution. Carriage speed, and therefore inflow velocity was measured
using a linear encoder with 10 $\mu$m resolution. All of these
performance-related quantities were sampled at 2 kHz, while the tow tank's
motion controller provided redundant measurements of the carriage speed \textbf{\textit{[need to say how well the agreement between these two measurements are and which one is used throughout the paper; else, remove one of these]}} turbine angular velocity sampled at 1 kHz. Turbine wake measurements at 1 meter
downstream were measured with a Nortek Vectrino+ acoustic Doppler velocimeter,
sampling at 200 Hz. A photo and drawing of the experimental setup are shown in
Fig.~\ref{fig:exp-setup-photo} and Fig.~\ref{fig:exp-setup}, respectively.
